### 2018-03-25 paper/paper.tex
intro and stuff...

% Koma-Script Basisklasse
\documentclass[a4paper,12pt,pagesize,headsepline,bibtotoc,titlepage,abstracton]{scrartcl}

\usepackage[utf8]{inputenc}		% direkte Eingabe von Umlauten & Co. (Vorsicht: Encoding im Editor muss auch UTF-8 sein!)

\usepackage[T1]{fontenc}			% T1-Schriften

\usepackage{mathptmx}			% Times/Mathe \rmdefault
\usepackage[scaled=.90]{helvet}	% Skalierte Helvetica \sfdefault
\usepackage{courier}			% Courier \ttdefault

% Zusatzpakete für mehr mathematische Symbole, Einfügen von Grafiken 
% und bessere Bildunterschriften
\usepackage{amsmath,amsthm,amsfonts,graphicx,caption}

% Wenn man direkt mit dem pdflatex eine PDF-Datei erzeugt, sollten diese beiden Pakete eingebunden werden
\usepackage{hyperref} % Hyperlinks anklickbar
\usepackage{ae,aecompl} % bessere Bildschirmschriftarten usw.
\usepackage{epstopdf} % support eps

\pagestyle{headings}

% Abstand der Kopfzeile vom Text:
\headsep4mm

\typearea[current]{current}     % Satzspiegel neu berechnen

% andere Bildunterschrift mit Hilfe von caption
\renewcommand{\figurename}{Fig.}
\renewcommand{\captionlabelfont}{\bf}

\newcommand{\quot}[1]{{``#1''}}

\usepackage{color}

\title{
	\includegraphics*[width=0.4\textwidth]{hpi_logo_2017.eps}\\
	\vspace{24pt}
	Genetic Doodle Jump
}
\subtitle{
	Seminar\\
	Machine Intelligence with Deep Learning\\
	Fall Semester 2017/2018
}
\author{
	Marcel Jankrift\\
	Fabian Sommer\\
	Toni Stachewicz\\[12pt]
	Supervisor:\\
	Goncalo Mordido,
	Dr. Haojin Yang\\
	Prof. Dr. Christoph Meinel
}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

\newpage

\begin{abstract}
Games usually have a wide range of possible parameters to tune. Often times, simple neural networks can meet all the requirements and perform well. However, training the net often takes many hours or a few days. The training time can be reduced tremendously by using neural nets in combination with a genetic algorithm. In this paper we present how we applied a deep genetic algorithm on the game \quot{Doodle Jump}. We especially pay attention to the configuration of the genetic algorithm, describe the undertaken changes and show, which influence they have on the performance.
\vskip 0.7cm
\noindent \textbf{Keywords.} Deep Learning, Genetic Algorithm, Artificial Intelligence, Games
\end{abstract}
\newpage

\section{Introduction}
Teaching an AI to play a simple game has become an interesting research topic in the last couple of years. AlphaGo\footnote{\url{https://deepmind.com/research/alphago/}} by Google Deepmind is probably the most famous example of computers playing games. The algorithm plays the board game Go, whose complexity is orders of magnitude higher than the complexity of chess. In a tournament against Lee Sedol (one of the world's best Go player) in 2016 the AI won four out of five games.

We did something similar in our research seminar, but with a less complex game: Doodle Jump. The player of this video game has to jump on randomly located platforms in order to get higher and thus, get a higher score. In our approach we use a genetic algorithm in combination with neural networks ??/for the machine learning part. An artificial neural net (ANN) is a machine learning algorithm inspired by nature. It consists of neurons, which calculate an output based on an input similar to a biological nerve cell. Those neurons are connected with each other to enable an information flow within the network. In order to improve the performance of an ANN, many parameters --- for example the biases of the neurons or the weight of connections --- can be tuned.

At this point we make use of the genetic algorithm. This algorithm is modelled on biological evolution. There are many individuals which form a population. Applied to our implementation, those are the different neural nets playing the game. After all players died, we select the best of them and perform crossover and mutation like in nature on those. Hence, we create a new generation, which is supposed to be better than the previous one. In this way, we always create better and better neural networks.

In the next two sections, we introduce the basics of the Doodle Jump game in detail and give a rough overview of our user interface. After that, we pay special attention to the used neural network. In section \ref{sec:ga} we present our genetic algorithm, in particular with focus on the basic operations of this algorithm and its evalution. Finally, we conclude our work and show some interesting findings.

\section{Doodle Jump}
\subsection{Original game}

The video game Doodle Jump is developed for mobile devices by Lima Sky\footnote{\url{http://www.limasky.com/}}. The player --- the so called \quot{doodler} --- has to jump on platforms in order to get a higher score. There are four different types of platforms:
\begin{enumerate}
    \item Regular ones, which neither move nor break
    \item Horizontally moving platforms
    \item Vertically moving platforms
    \item Vanishable ones, which disappear after jumping on it once
    \item Breaking platforms, where the doodler can not jump on
\end{enumerate}
The platforms are pseudo randomly located on the map. Pseudo randomly located means, that there will always be at least one platform, which is reachable from the previous one and non-breakable (See Fig. \ref{abb:doodlejumpgame}). In other words, there is always a way to reach an infinite score. However, the higher the achieved score, the more difficult the game becomes. The difficulty is reached via higher distances between the platforms and the increased usage of moving and vanishable platforms.

The game also includes bonus features, such as propeller hats, jetpacks, rockets, springs or trampolines. These allow the player to jump higher and faster and therefore skip some platforms. Indeed, there are also \quot{enemies} in the game for example monsters, UFOs or black holes. If the doodler hits a monster or comes close to a black hole or UFO, it dies and the game is over. Monsters and UFOs can also be killed by shooting or jumping on them.

To control the doodler the player of the game has to tilt his mobile devices left and right. If the doodler hits the left or right boundary of the game, it will appear on each other's side. Shooting is performed by tapping on the screen. The downrange depends on the tapping position. Once a player falls down and hits the bottom the game is over.

\begin{figure}[h]
\begin{center}
\includegraphics*[width=0.4\textwidth]{images/game}\\
\caption{Screenshot of original Doodle Jump game}
\label{abb:doodlejumpgame}
\end{center}
\end{figure}

\subsection{Simplified implementation}

Since we do not want to perform the deep learning part on a mobile device or within an emulator, we decided to use an implementation, which runs in a browser. We are using \quot{HTML5 Doodle Jump} by ntcnet83\footnote{\url{https://github.com/ntcnet83/html5-doodle-jump}}, which we found on github\footnote{\url{https://github.com/}}. This custom version of the is written in HTML5 and JavaScript. The doodler can be control via left and right arrow keys. In contrast to the original implementation, this version of the game contains less features, which makes the game simpler at the end:
\begin{itemize}
\item Except springs, there are no other bonuses.
\item There are no enemies. Hence, the only way to die is falling down.
\item The feature shooting is not necessary and therefore not available.
\item The game does not have vertically moving platforms.
\item The difficulty changes after a certain score, but only through the usage of more moving and vanishable platforms. The distances between the platforms stay the same.
\end{itemize}

Our machine learning part runs in the browser as well. We use the \textit{Synaptic} library\footnote{\url{https://github.com/cazala/synaptic}} in order to create our neural networks. This library provides a high level interface to simply initialize a perceptron. Furthermore, adapted the genetic algorithm implemented in the \textit{Genetic.js} library\footnote{\url{https://github.com/subprotocol/genetic-js}}. Both libraries are written in JavaScript.

\section{User Interface}

\section{Network Layout}

\section{Genetic algorithm}
\label{sec:ga}

Whether or not the neural network performs well at playing the game very heavily depends on the specific parameters it uses: Each node possesses a bias and each edge has a weight. Even though the network is rather small, there are too many weights and biases to optimize them manually.

We therefore employ a genetic algorithm to obtain good parameters for the network. At any time, there is a number of different networks, so-called candidates, all playing the game independent of each other. Once they are all done, their performance is evaluated using a fitness function. The best candidates are then selected and combined in different ways to obtain the next generation of candidates. This combination process is called crossover. Finally, some parameters of some candidates are altered (mutated) to make it possible to generate any network. This new generation is then evaluated again, and the process is iterated.

Doodle Jump is a game where each level is generated randomly. We want to end up with a network that is able to play any level well, regardless of the specific generated layout. The performance of any one network on any specific level may not necessarily be indicative of the networks ability to play the game overall. Therefore, each network plays several levels simultaneously, and each level is taken into account when evaluating the performance.

In order to make the performance of different candidates comparable, all candidates of a generation play the same set of levels. However, for each generation a new set of levels is generated. Otherwise, we might end up with a network that excels in playing the specific set of levels used for evaluation but fails at any other level. Therefore, we can avoid overfitting.

We want to choose the different parts of the genetic algorithm so that the learning process is as quick as possible, i.e. the average score achieved by the candidates after 100 generations should be as high as possible.

\subsection{Fitness function}

The fitness function measures the ability of the candidates to play the game well. Doodle Jump already provides a score which measures the height a player achieves. Since the games aim is to maximize that score, we used the score as initial fitness function.

Beyond that, there are many other fitness function one could think of. For example, at the very beginning of the optimization process, many candidates might do nothing at all, while others move a little bit into the direction of the next platform, but not far enough to actually land on it. All candidates would thus achieve a score of zero, and the primitive fitness function would be unable to discriminate between them. Another fitness function, which rewards candidates that got closer to the platform they failed to jump on might speed up the learning process in the first few generations. It would also be helpful later on, when different candidates might fail at the same platform, thus achieving an equal score. However, such a fitness function would include prior knowledge about how to play the game well. We want to avoid using such fitness functions, since for other problems, prior knowledge might not be available, or might not lead to the optimal solution if available.

One measurement that does not involve any prior knowledge is time. Between two candidates, both achieving similar score, where one jumps once on every platform and the other has to jump twice on every platform before attempting the next jump, we want to prefer the first candidate since he is climbing twice as fast. Therefore, the required time should negatively affect the fitness of candidates.

We tested multiple fitness functions achieving this in different ways:
\begin{itemize}
\item $fitness = score$
\item $fitness = \frac{score}{time}$
\item $fitness = score - time$
\item $fitness = normalized(score) - normalized(time)$
\item $fitness = score^2 - time$
\item $fitness = normalized(score^2 - time)$
\end{itemize}

For each fitness function, we monitored the learning process for the first 100 generations. Since it is a random process with a high variance, we repeated this TODO times for each function. Figure TODO shows the test results.

\subsection{Selection}
\subsection{Crossover}
\subsection{Mutation}
\subsection{Retaining best candidates}

\section{Interesting findings}
\subsection{Disabling passing through walls}
Very good, simple strategy: Always go right -> local optimum.

When disabled and with (score - time) fitness function: Optimum where players suicide as quickly as possible. 
\subsection{Dealing with moving platforms}
\subsection{Dealing with fake platforms}
\subsection{Dealing with springs}
\section{Conclusion and Future Work}

\newpage
%%%% bibtex file %%%%%%
\bibliographystyle{alphadin}
\bibliography{references}

\end{document}